Effects of sexual selection on non-sexual fitness components: a meta-analysis

### 2018-03-02 meta_analysis_sexual_selection_pop_fitness.Rmd
2/3 Edit

**Table S3:** Table of effect sizes included in our meta-analysis. See the text following the data table for an explanation of each column.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r message=FALSE}
n.blind.ones <- (sum(prelim.data$Blind == "Blind"))
prelim.data %>% 
  summarise(
    Effect_sizes_.Totalq = n(), 
    Publications = prelim.data$Study.ID %>% unique() %>% length(),
    Blind_experiments = n.blind.ones,
    Effect_sizes_.Stressedq = (sum(prelim.data$Environment == "Stressed")),
    Effect_sizes_.Unstressedq = (sum(prelim.data$Environment == "Unstressed")),         
    Effect_sizes_.Maleq = (sum(prelim.data$Sex == "M")),
    Effect_sizes_.Femaleq = (sum(prelim.data$Sex == "F")),
    Effect_sizes_.Both_sexesq = (sum(prelim.data$Sex == "B")),
    Different_species =  prelim.data$Species %>% unique() %>% length(),
    Effect_sizes_.Beetleq = sum(Taxon == "Beetle"),
    Effect_sizes_.Flyq = sum(Taxon == "Fly"),
    Effect_sizes_.Mouseq = sum(Taxon == "Mouse"),
    Effect_sizes_.Nematodeq = sum(Taxon == "Nematode"),
    Effect_sizes_.Miteq = sum(Taxon == "Mite"),
    Effect_sizes_.Cricketq = sum(Taxon == "Cricket"),
    Effect_sizes_.Guppyq = sum(Taxon == "Guppy")) %>% melt() %>%
  mutate(variable = gsub("_", " ", variable),
         variable = gsub("[.]", "(", variable),
         variable = gsub("q", ")", variable)) %>% 
  rename_("n" = "value", " " = "variable") %>% 
  pander(split.cell = 40, split.table = Inf)
```
 <br/><br/> 
**Table S4:** Table of fitness outcomes included in our meta-analysis by sex.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
Outcome_and_sex <- table(prelim.data$Outcome, prelim.data$Sex) %>%
  pander(split.cell = 40, split.table = Inf)
```

@@

#Generate a plot
p.meta <- prelim.data %>% 
  mutate(Sex = replace(as.character(Sex), Sex == "B", "Both"),
         Sex = replace(Sex, Sex == "M", "Male"),
         Sex = replace(Sex, Sex == "F", "Female")) %>%
  ggplot(aes(y=reorder(AuthorYear, -g), x = g, shape = Sex, color = Outcome.Class)) +
  geom_errorbarh(aes(xmin = lowerci, 
                     xmax = upperci), height = 0.1) +
  geom_point(size = 3) +  #Add data points and color them black
  scale_x_continuous(limits=c(-5, 5), name='Standardized Mean Difference (g)') +
  scale_color_discrete(name = "Relationship\nto fitness") + 
  ylab('Reference') + 
  geom_vline(xintercept=0, color='black', linetype='dashed')+
  facet_grid(Outcome_f~., scales= 'free', space='free')+
  theme(strip.text.y = element_text(angle = 0))+
ggsave(plot = p.meta, filename = "figures/Big_forest_plot.eps", height = 25, width = 12)
p.meta
```
<br></br>
**Figure 1:** Forest plot of raw effect sizes and their 95% confidence intervals, grouped according to environment (stressed or unstressed) and the sex of the individuals whose fitness trait was measured (male, female, or both sexes mixed together). Rows with multiple data points denote studies that provided multiple effect sizes.

@@

summary(model.preliminary)

```
<br></br>
**A note on random effects** Here we utilise crossed random effects of Study.ID and Outcome. Nesting Outcomes within Study.ID can also be achieved through ``random = ~ 1 | Study.ID/Outcome``. However given that the same outcome is expected to be measured in the same/very similar way between studies we used a crossed random effects design.

@@

```{r, fig.height= 5, fig.width=7.5}
pd <- position_dodgev(height = .7)
Taxon.metaanlysis <- predictions.complete %>% ggplot(aes(x = pred, 
                                                         y= Environment, 
                                                         fill = Sex)) + 
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70") + 
  geom_errorbarh(aes(xmin = predictions.complete$ci.lb, 
                     xmax = predictions.complete$ci.ub,
                     color= Sex), 
                 height = 0, position = pd) +
  geom_point(position = pd, size=3, shape=21, color = "grey20") + 
  facet_grid(Taxon ~.)+
  ylab("Environment \n")+
  xlab("\nModel Prediction (Hedges g)")+
  xlim(-2, 2)+
  ggtitle('Effects of Sex, Stress and Outcome Class \non Population Fitness for Each Taxon')+
  scale_fill_brewer(palette = "Set1")+
  scale_color_brewer(palette = "Set1")

ggsave(plot = Taxon.metaanlysis, filename = "figures/Taxon_metaanalysis_plot.eps", height = 5, width = 7.5)
Taxon.metaanlysis
```

@@

```{r, fig.height= 5, fig.width=7.5}

@@

#And plot the results
pd <- position_dodgev(height = .5)
EnvSex.metaanalysis <- predictions.complete2 %>% ggplot(aes(x = pred, y= Environment, fill = Sex)) + 
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70") + 
  geom_errorbarh(aes(xmin = predictions.complete2$ci.lb, 
                     xmax = predictions.complete2$ci.ub,
                     color = Sex), height = 0, position = pd) +
  geom_point(position = pd, size=3, shape=21, color = "grey20") + 
  ylab("Environment\n")+
  xlab("\nEffect Size (Hedges g)")+
  xlim(-.75, .75)+
  ggtitle('Effects of Sex, Stress \non Population Fitness')+
  scale_fill_brewer(palette = "Set1")+
  scale_color_brewer(palette = "Set1")

ggsave(plot = EnvSex.metaanalysis, filename = "figures/Environment_Sex_metaanalysis_plot.eps", height = 5, width = 7.5)
EnvSex.metaanalysis
```

@@

ggsave(plot = var.plot, filename = "figures/Variance_plot.eps", height = 5, width = 7.5)
var.plot
```

@@

```{r, fig.height= 7, fig.width=10}
#Funnel plot for all outcome classes

@@

  ggplot(dataset, aes(x = sqrt(var.g), y = g)) +
    geom_point(alpha = 0.4) +
    xlab("Standard Error") + ylab("Effect size (Hedges' g)") +
    geom_line(aes(x = se.seq, y = ll95), linetype = 'dotted', data = dfCI) + # confidence lines
    geom_line(aes(x = se.seq, y = ul95), linetype = 'dotted', data = dfCI) +
    geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
    geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
    #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
    geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI, colour = "tomato") +
    geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI, colour = "tomato") +
    scale_x_reverse() +
    # scale_y_continuous(breaks = seq(-1.25,2,0.25)) + #Choose values that work for you based on your data
    coord_flip() +
    theme_classic()
    # apatheme

@@

#make.funnel(restricted.data, model.complete2)
```

@@

```{r, fig.width=5, fig.height=7.5}
prelim.data %>% ggplot(aes(x=log10(JIF), y=(g)))+
  geom_jitter(color='darkgreen', alpha=.3, aes(size = (1/(var.g))))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'weight')
```

> I think it may also be interesting to plot the maximum effect size for each study and then plot the results as many studies measure a suite of traits.

@@

```{r, fig.width=5, fig.height=7.5}
prelim.data %>% 
  ggplot(aes(x=Year, y=g))+
  geom_jitter(color='orange', alpha=.3, aes(size = (1/(var.g))))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'weight')
```

@@

From these graphs, we see a small trend for larger effect sizes to be published in higher impact journals as well as for effect size to decrease over successive years. Additionally to publication bias, other forms of bias may exist within studies. We initially collected data on whether studies were blind or not. Although not enough studies were blind for us to include it as a fixed effect within the model we can test whether blinding affects the raw effect size:

```{r, fig.width=5, fig.height=7.5}
df.forest.model %>% ggplot(aes(x=Blinding, y=g))+
  geom_boxplot()+
  geom_jitter(aes(color=Blinding))+
  geom_hline(yintercept=0, linetype = 'dotted') + 
  theme(legend.position = "none")
```

@@

We also collected sample sizes for each of the effect sizes calculated. Because we are dealing withj different taxa some studies are not suited to have sample sizes in the 1000's. We can simply inspect the sample size ande effect sizes through the following plot:
```{r, fig.width=5, fig.height=10}
prelim.data %>% ggplot(aes(x=n, y = g))+
  geom_point()+
  scale_x_log10(breaks = c(10, 100, 1000, 10000)) + 
  #xlim(0,2100)+
  facet_grid(Taxon~., scales='free')+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed")
```

(Promislow 1998 has one sample size of >10,000) and is not shown here. From these plots we can see that with increased sample size the effect sizes are closer to zero. Thistrend should be taken into account as meta-analytic models are wighted by 1/variance.

@@

```{r, fig.width=8, fig.height=7}
prelim.data %>% ggplot(aes(x=Generations, y=g, fill=Taxon))+
  geom_point(shape=21, color = "grey20")+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed") + 
  scale_fill_brewer(palette = "Set3")
```

>>>>

**Table S3:** Table of effect sizes included in our meta-analysis. See the text following the data table for an explanation of each column.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r message=FALSE}
n.blind.ones <- (sum(prelim.data$Blind == "Blind"))
prelim.data %>% 
  summarise(
    Effect_sizes_.Totalq = n(), 
    Publications = prelim.data$Study.ID %>% unique() %>% length(),
    Blind_experiments = n.blind.ones,
    Effect_sizes_.Stressedq = (sum(prelim.data$Environment == "Stressed")),
    Effect_sizes_.Unstressedq = (sum(prelim.data$Environment == "Unstressed")),         
    Effect_sizes_.Maleq = (sum(prelim.data$Sex == "M")),
    Effect_sizes_.Femaleq = (sum(prelim.data$Sex == "F")),
    Effect_sizes_.Both_sexesq = (sum(prelim.data$Sex == "B")),
    Different_species =  prelim.data$Species %>% unique() %>% length(),
    Effect_sizes_.Beetleq = sum(Taxon == "Beetle"),
    Effect_sizes_.Flyq = sum(Taxon == "Fly"),
    Effect_sizes_.Mouseq = sum(Taxon == "Mouse"),
    Effect_sizes_.Nematodeq = sum(Taxon == "Nematode"),
    Effect_sizes_.Miteq = sum(Taxon == "Mite"),
    Effect_sizes_.Cricketq = sum(Taxon == "Cricket"),
    Effect_sizes_.Guppyq = sum(Taxon == "Guppy")) %>% melt() %>%
  mutate(variable = gsub("_", " ", variable),
         variable = gsub("[.]", "(", variable),
         variable = gsub("q", ")", variable)) %>% 
  rename_("n" = "value", " " = "variable") %>% 
  pander(split.cell = 40, split.table = Inf)
```
 <br/><br/> 
**Table S4:** Table of fitness outcomes included in our meta-analysis by sex.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
Outcome_and_sex <- as.data.frame.matrix(table(prelim.data$Outcome, prelim.data$Sex))
colnames(Outcome_and_sex) <- cbind("Both", "Male", "Female")

Outcome_and_sex %>% pander(split.cell = 40, split.table = Inf)

@@

#Generate a plot
p.meta <- prelim.data %>% 
  mutate(Sex = replace(as.character(Sex), Sex == "B", "Both"),
         Sex = replace(Sex, Sex == "M", "Male"),
         Sex = replace(Sex, Sex == "F", "Female")) %>%
  ggplot(aes(y=reorder(AuthorYear, -g), x = g, shape = Sex, color = Outcome.Class)) +
  geom_errorbarh(aes(xmin = lowerci, 
                     xmax = upperci), height = 0.1) +
  geom_point(size = 3) +  #Add data points and color them black
  scale_x_continuous(limits=c(-5, 5), name='Standardized Mean Difference (g)') +
  scale_color_discrete(name = "Relationship\nto fitness") + 
  ylab('Reference') + 
  geom_vline(xintercept=0, color='black', linetype='dashed')+
  facet_grid(Outcome_f~., scales= 'free', space='free')+
  theme(strip.text.y = element_text(angle = 0))
ggsave(plot = p.meta, filename = "figures/Big_forest_plot.eps", height = 25, width = 12)
p.meta
```
<br></br>
**Figure 1:** Forest plot of raw effect sizes and their 95% confidence intervals, grouped according to environment (stressed or unstressed) and the sex of the individuals whose fitness trait was measured (male, female, or both sexes mixed together). Rows with multiple data points denote studies that provided multiple effect sizes.

@@

summary(model.preliminary)
```
<br></br>
**A note on random effects** Here we utilise crossed random effects of Study.ID and Outcome. Nesting Outcomes within Study.ID can also be achieved through ``random = ~ 1 | Study.ID/Outcome``. However given that the same outcome is expected to be measured in the same/very similar way between studies we used a crossed random effects design.

@@

```{r, fig.height= 7.5, fig.width=10}
pd <- position_dodgev(height = .7)
Taxon.metaanlysis <- predictions.complete %>% ggplot(aes(x = pred, 
                                                         y= Environment, 
                                                         fill = Sex)) + 
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70") + 
  geom_errorbarh(aes(xmin = predictions.complete$ci.lb, 
                     xmax = predictions.complete$ci.ub,
                     color= Sex), 
                 height = 0, position = pd) +
  geom_point(position = pd, size=3, shape=21, color = "grey20") + 
  facet_grid(Taxon ~.)+
  ylab("Environment \n")+
  xlab("\nModel Prediction (Hedges g)")+
  xlim(-2, 2)+
  ggtitle('Effects of Sex and Stress on \nPopulation Fitness for Each Taxon')+
  scale_fill_brewer(palette = "Set1")+
  scale_color_brewer(palette = "Set1")

ggsave(plot = Taxon.metaanlysis, filename = "figures/Taxon_metaanalysis_plot.eps", height = 7.5, width = 10)
Taxon.metaanlysis
```

@@

```{r, fig.height= 7.5, fig.width=10}

@@

#And plot the results
pd <- position_dodgev(height = .5)
EnvSex.metaanalysis <- predictions.complete2 %>% ggplot(aes(x = pred, y= Environment, fill = Sex)) + 
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70") + 
  geom_errorbarh(aes(xmin = predictions.complete2$ci.lb, 
                     xmax = predictions.complete2$ci.ub,
                     color = Sex), height = 0, position = pd) +
  geom_point(position = pd, size=3, shape=21, color = "grey20") + 
  ylab("Environment\n")+
  xlab("\nEffect Size (Hedges g)")+
  xlim(-.75, .75)+
  ggtitle('Effects of Sex and Stress \non Population Fitness')+
  scale_fill_brewer(palette = "Set1")+
  scale_color_brewer(palette = "Set1")

ggsave(plot = EnvSex.metaanalysis, filename = "figures/Environment_Sex_metaanalysis_plot.eps", height = 7.5, width = 10)
EnvSex.metaanalysis
```

@@

ggsave(plot = var.plot, filename = "figures/Variance_plot.eps", height = 7.5, width = 10)
var.plot
```

@@

```{r, fig.height= 7.5, fig.width=10}

#Using residuals for the funnel plot means that we need to generate residuals (intercept only)

forest.model <- rma.mv(g, var.g,
                       mods = ~ 1,
                       random = list(~ 1 | Study.ID,
                                       ~ 1 | Outcome),
                       method = "REML",
                       data = prelim.data)

# Obtain residuals
resstandards <- (rstandard.rma.mv(forest.model,
                                   type="response"))

# Obtain grand mean effect size    <- ADDED BY LUKE
grand.mean <- as.numeric(forest.model$b)

# Create new df with residuals replacing raw
df.forest.model <- prelim.data
df.forest.model$g <- resstandards$resid + grand.mean 
df.forest.model$sei <- resstandards$se

# Funnel plot for all outcome classes

@@

  ggplot(dataset, aes(x = sqrt(var.g), y = g)) +
    geom_point(size=1, shape = 21, color= "grey20") +
    xlab("Standard Error") + ylab("Effect size (Hedges' g)") +
    geom_line(aes(x = se.seq, y = ll95), linetype = 'dotted', data = dfCI) + # confidence lines
    geom_line(aes(x = se.seq, y = ul95), linetype = 'dotted', data = dfCI) +
    geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
    geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
    #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
    geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI, colour = "tomato") +
    geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI, colour = "tomato") +
    scale_x_reverse() +
    # scale_y_continuous(breaks = seq(-1.25,2,0.25)) + #Choose values that work for you based on your data
    coord_flip() +
    scale_fill_brewer(palette = "Set1")+
    theme_classic()
    # apatheme

@@

funnel.plot <- make.funnel(df.forest.model, forest.model)

ggsave(plot = funnel.plot, filename = "figures/funnel_plot.eps", height = 7.5, width = 10)
funnel.plot

@@

```{r, fig.height= 7.5, fig.width=10}
prelim.data %>% ggplot(aes(x=JIF, y=g))+
  geom_jitter(color='darkgreen', alpha=.3, aes(size = (1/(var.g))/sum((1/(var.g)))*100))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'Weight (%)', y='Effect size (Hedges g)', x= 'Journal Impact Factor')
```

> Is this graph ok how it is?

@@

```{r, fig.height= 7.5, fig.width=10}
prelim.data %>% 
  ggplot(aes(x=Year, y=g))+
  geom_jitter(color='darkorange', alpha=.5, aes(size = (1/(var.g))/sum((1/(var.g)))*100))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'Weight (%)', y='Effect size (Hedges g)', x= 'Year')
```

@@

From these graphs, we see a small trend for larger effect sizes to be published in higher impact journals as well as for effect size to decrease over successive years. Additionally to publication bias, other forms of bias may exist within studies. We initially collected data on whether studies were blind or not. Although not enough studies (n=8) used blinding for us to include it as a fixed effect within the model we can test whether blinding affects the raw effect size:

```{r, fig.height= 7.5, fig.width=10}
df.forest.model %>% ggplot(aes(x=Blinding, y=g))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(fill=Blinding), shape=21, color='grey20', size=1.7)+
  geom_hline(yintercept=0, linetype = 'dotted') + 
  theme(legend.position = "none")+
  scale_fill_brewer(palette = "Set2")+
  labs(y='Effect size (Hedges g)', x= 'Blinding')
```

@@

We also collected sample sizes for each of the effect sizes calculated. Because we are dealing withj different taxa some studies are not suited to have sample sizes in the 1000's. We can simply inspect the sample size ande effect sizes through the following plot:

```{r, fig.height= 10, fig.width=5}
prelim.data %>% ggplot(aes(x=n, y = g))+
  geom_point(fill="grey", color="grey20", shape=21)+
  scale_x_log10(breaks = c(10, 100, 1000, 10000)) + 
  #xlim(0,2100)+
  facet_grid(Taxon~., scales='free')+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed")+
  labs(y='Effect size (Hedges g)', x= 'Sample Size (n) [logscale]')
```

(Promislow 1998 has one sample size of >10,000) and is not shown here. From these plots we can see that with increased sample size the effect sizes are closer to zero. This trend should be taken into account as meta-analytic models are wighted by 1/variance.

@@

```{r, fig.height= 7.5, fig.width=10}
prelim.data %>% ggplot(aes(x=Generations, y=g, fill=Taxon))+
  geom_point(shape=21, color = "grey20", size=2)+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed") + 
  scale_fill_brewer(palette = "Set3")+
  labs(y='Effect size (Hedges g)', x= 'Generations')
```
